Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-21 Chapter-2/2-LArTPCs.tex
s

\subsubsection{Purity \& Electron Life Time }
The presence of electronegative contaminants in liquid argon, such as oxygen and water, is particularly
pernicious, since these molecules quench the charge produced by the ionizing radiation.  Thus, amount of charge per unit of length $dQ/dx$ collected on the collection plane depends on the charge's production point in the detector: a ionization produced  close to the cathode will see more impurities along its journey to the collection plane than ionization produced close to the anode, resulting in greater attenuation. As a result,  the amount of charge collected on the sense wires as a function of the traveled distance follows an exponential decay trend. The traveled distance is generally measured in terms of drift time and the  characteristic time constant of the exponential decay is called electron lifetime $\tau_e$. Figure \ref{fig:Elifetime} shows the typical life time for LArIAT data \textcolor{red}{cite technote}. LArIAT small drift distance (47 cm) allows for a relatively short electron life time. The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of  to allow charge collection usable for physics analyses. Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section s\ref{ch:energyCalibration}.

@@

\subsubsection{Space Charge Effect}
\subsubsection{Recombination Effect}

>>>>

\subsubsection{Purity \& Electron Life Time }
The presence of electronegative contaminants in liquid argon, such as oxygen and water, is particularly
pernicious, since these molecules quench the charge produced by the ionizing radiation.  Thus, amount of charge per unit of length $dQ/dx$ collected on the collection plane depends on the charge's production point in the detector: a ionization produced  close to the cathode will see more impurities along its journey to the collection plane than ionization produced close to the anode, resulting in greater attenuation. As a result,  the amount of charge collected on the sense wires as a function of the traveled distance follows an exponential decay trend. The traveled distance is generally measured in terms of drift time and the  characteristic time constant of the exponential decay is called electron lifetime $\tau_e$. Figure \ref{fig:Elifetime} shows the typical life time for LArIAT data \textcolor{red}{cite technote}. LArIAT small drift distance (47 cm) allows for a relatively short electron life time. The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of  to allow charge collection usable for physics analyses. Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section \ref{ch:energyCalibration}.

@@

\subsubsection{Recombination Effect}
After production, ionization electrons thermalize with the surrounding medium and may recombine with nearby ions. Recombination might occur either between the electron and the parent ion through Coulomb attraction (Onsager geminate theory  [2] ) or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory (Jaffe[3]  columnar model ). 
Consideration on the  average electron-ion distance and the average ion-ion distance for argon show that the probability of geminate recombination is low; thus recombination in argon is mainly due to collective effects\cite{1748-0221-8-08-P08005}.  Since protons, kaons and stopping particles present a higher ionization compared to MIPs, recombination effects are more prominent when considering the reconstruction of energy deposited by these particles.

Models for a theoretical descriptions of recombination base on are provided in (Birks model [6]) and in (Box model [7]). The Birks model assumes a gaussian spatial distribution around the particle trajectory during the entire recombination phase and identical charge mobility for ions and electrons. The Box model also assumes that electron diffusion and ion mobility are negligible in liquid argon during recombination
In these models, the fraction of ionization electrons surviving recombination is a function of the number of ion-electron pairs per unit length, the electric field, the average ion-electron separation distance after thermalization and the angle of the particle with respect to the direction of the electric field -- plus the diffusion coefficient in the Birks model. Given the stringent assumptions, it is perhaps  not surprising that these models are in accordance to data only in specific regimes: the Birks model is generally used to describe recombination for low dE/dx, the Box model for high dEdX.
In LArTPC, the ICARUS and ArgoNeut have measured recombination in [8] and \cite{1748-0221-8-08-P08005} respetively. Since LArIAT uses the refurbished ArgoNeut TPC and cryostat,  LArIAT currently corrects for recombination using the ArgoNeut measurement, shown in figure \textcolor{red}{find figure}.

\subsubsection{Space Charge Effect}
Slow-moving positive argon ions created during ionization can build-up in LArTPC, causing the distortion of the
electric field within the detector. This effect, called  ``space charge effect" leads to a displacement in the reconstructed position of the signal ionization electrons. In surface LArTPCs the space charge effect is primarily due to the rate of ionization produced by cosmic rays which is slowly drifting in the chamber at all times. Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm. The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to b even smaller. \textcolor{red}{CHIEDI A FLAVIO}
